---
title: "Template 05_2020"
author: "Veronica Thompson"
date: "May 1, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("lme4")
library("lmerTest")
library(emmeans)
```

## Project title
### project subtitle
### project data

### Input aesthectics  
set to blue/green qualitative palette for now
```{r aesthetics}
#edit to rename genotypes on the plots
Tem_name_key = tibble(
  geno = c("WT", "elf3"),
  new_geno = c("WT", "elf3")
)

#set level and colors
Tem_levels = c("GV342", "elf3-1")
Tem_colors = c("#1f78b4", "#33a02c")
```

### Relative Expression plots  
load data  
```{r}
#load  
expression_data <- read_csv("master_flowering_time_data_2020_05_01.csv")
                     col_types = cols(
                       genotype = col_factor(),
                       timepoint = col_factor(),
                       sample = col_factor(), 
                       GOI = col_factor()))

#rename genotypes  
expression_data = left_join(expression_data, geno_key, by = "genotype")

#relevel genotypes  
expression_data$new_geno = factor(expression_data$new_geno, levels = new_levels)

#calculate SD points
expression_data = expression_data %>%
  mutate(
    error_low = rel_exp - stand_dev,
    error_high = rel_exp + stand_dev
  )
  
expression_data
```

Plot
```{r}
ggplot(expression_data, aes(new_geno, rel_exp, fill = new_geno, linetype = timepoint)) + 
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 1) + #size changes line weight
  geom_errorbar(aes(ymin = error_low, ymax = error_high),
                position = "dodge", size = 0.8) +
  facet_grid(. ~ GOI) +
  labs(title = "Relative expression levels",
       subtitle = "elf3 complementation T2 families",
       x = "genotype",
       y = "relative expression/AtPP2A",
       fill = "Genotype") +
  theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = color_key)
```

## Hypocotyl data  

load and check data:
```{r}
hyp_data <- read_csv("elf3_T2_hypocotyl_measurements_SC_04_2020_compiled.csv",
                     col_types = cols( 
                       Plate_Number = col_factor(),
                       Genotype = col_factor()))

#transform hyp to mm
hyp_data$Hyp_Length = hyp_data$Hyp_Length*10

#rename Genotype to genotype so our new key works
names(hyp_data)[5] <- "genotype"

hyp_data = left_join(hyp_data, geno_key, by = "genotype")

#relevel genotypes
hyp_data$new_geno = factor(hyp_data$new_geno, levels = new_levels)

#one of the elf3 points is way too short, remove that.
hyp_data = filter(hyp_data, No. != 130)

summary(hyp_data)
```

plot  
```{r}
ggplot(hyp_data, aes(new_geno, Hyp_Length, fill = new_geno)) + 
  geom_boxplot(outlier.size = -1) + 
  geom_point(alpha = 0.5, position = position_jitterdodge()) +
  labs(title = "Hypocotyl elongation",
       subtitle = "elf3 complementation T2 families",
       x = "genotype",
       y = "hypocotyl length (mm)",
       fill = "Genotype") +
  scale_fill_manual(values = color_key) + 
  scale_y_continuous(breaks = c(1:9))
```

## Stats walk-through  
### This was helpful for my refernece but not useful for plotting in the end  
from lab 3 of MCB 205 - 2018  
From Dan Runcie  

Steps:  
1. Visualize the data - done above  
2. Fit a model to the data  
3. Run model diagnostics to check model assumptions  
4. Use the model fit to calculate summary statistics and draw conclusions  

```
hyp_model = lmer(Hyp_Length ~ Genotype + (1|Plate_Number), data = hyp_data)
hyp_model
```  


```
plot(hyp_model)
```   

residuals vs. fitted plot. Higher residuals for larger fitted values isn't great, but this is workable.  

```
means_hyp = emmeans(hyp_model,~Genotype)  #preps emmeans analysis, group by geno
summary(means_hyp,level = 0.95,infer = c(T,F))  # note setting level (1-alpha) for the confidence intervals
# infer = c(T,F) gives confidence intervals
# infer = c(F,T) gives p-values intervals. Here these are testing against the null hypothesis that mu_i = 0
# infer = c(T,T) gives both.
```  

```
differences_geno = emmeans(hyp_model, trt.vs.ctrl ~ Genotype, ref = 2)
summary(differences_geno,level = 0.95,infer = c(T,T))$contrast
```  
pretty small differences between elf3 and T2 families but most are significant... 

compact letter display seems to have fallen out of favor. Is this the comparison I want to make (elf3 as control vs all others)? If so I should reorder my plot so elf3-1 is first...  