---
title: "week5_advanced"
author: "Veronica Thompson"
date: "5/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Week 5 advanced problems  

#### champter 19

19.2.1 #3  
19.4.4 either #2 or #3 (pick whichever you want)  

Question 19.2.1 #3  
Practice turning the following code snippets into functions. Think about what each function does. What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?  

```{r}
x <- c(1:10, NA) 

mean(is.na(x))
```

is.na() evaluates each as true or false, then mean() asseses mean such that 0 = F and 1 = T.  
only one input, x  
I don't think I can rewrite it?  
```{r}
TF_average <- function(x) {
  mean(is.na(x))
}

tibs = tibble(
  col_a = c(NA, NA, 10, 12),
  col_b = c(111:114),
  col_c = c(NA, NA, NA, 1)
)

TF_average(tibs$col_a)
TF_average(tibs$col_b)
TF_average(tibs$col_c)
```

```{r}
x / sum(x, na.rm = TRUE)
sum(x / sum(x, na.rm = TRUE), na.rm = T)
```

divides each value by the sum of all values in a vector, outputs proportion of total vector sum.  
```{r}
prop_of_total = function(x) {
  total = sum(x, na.rm = T) 
  x/total
}

prop_of_total(tibs$col_a)
prop_of_total(tibs$col_b)
prop_of_total(tibs$col_c)
```


```{r}
sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)

sd_over_mean = function(x) {
  input = c(x, na.rm = T)
  sd(input)/mean(input)
}

sd_over_mean(tibs$col_a)
sd_over_mean(tibs$col_b)
sd_over_mean(tibs$col_c)
```

Question 19.4.4 #3  
Implement a fizzbuzz function. It takes a single number as input. If the number is divisible by three, it returns “fizz”. If it’s divisible by five it returns “buzz”. If it’s divisible by three and five, it returns “fizzbuzz”. Otherwise, it returns the number. Make sure you first write working code before you create the function.  
```{r}
print(x)
x%%3 == 0
x%%5 == 0
```

code works!

```{r}
fizz_or_buzz_test <- function(x) {
  div_3 <- x%%3 == 0
  div_5 <- x%%5 == 0
  if (div_3) {
    "fizz"
  } else {"buzz"}
}

fizz_or_buzz_test(5)
fizz_or_buzz_test(3)
```

testing the base works, now add additional qualifiers.  

```{r}
fizz_or_buzz <- function(x, na.rm = T) {
  div_3 <- x%%3 == 0
  div_5 <- x%%5 == 0
  if (div_3) {  #div by three, then
    if(div_5) { #div by 3 and 5
      "fizzbuzz"
    } else{"fizz"} #div by 3 not 5
  } else if (div_5) {
    "buzz"
  } else(x)
}

fizz_or_buzz(15)
fizz_or_buzz(3)
fizz_or_buzz(5)
fizz_or_buzz(4)
fizz_or_buzz(NA)
```
 
 doesn't work with a vector, I think because of the logical statement. 
 
```{r}
fizz_or_buzz_vect <-Vectorize(fizz_or_buzz, vectorize.args = "x")

fizz_or_buzz_vect(tibs$col_b)
fizz_or_buzz_vect(tibs$col_c)
```
 
only works if there are no NAs. not sure how to correct this...


#### Chapter 21
21.2.1 #1 (pick 2 from the 4)  
21.3.5 #3  
21.5.3 #1 (pick the same 2 you did from 21.2.1)  

21.2.1 #1 
*1. compute the mean of every column in mtcars.*  
output should be a "double" vector
sequence is for i in mtcar columns
body is calculate mean
```{r}
output <- vector("double", ncol(mtcars))  #output
for (i in seq_along(mtcars)) {
  output[[i]] <- mean(mtcars[[i]])
}
output
mean(mtcars[,1])
mean(mtcars[,8])
```

looks good!  

*Compute the number of unique values in each column of iris* 
output should be a vector of "double" type, same length as number of iris columns.  
sequence is for i in iris columns  
body is count unique values in each column  
```{r}
length(unique(iris$Sepal.Length)) #length of vector for unique values

iris_unique_val <- vector("double", ncol(iris)) #vector of double, ncol(iris) long
for (i in seq_along(iris)) {
  iris_unique_val[[i]] <- length(unique(iris[[i]]))
}

iris_unique_val
```


21.3.5 #3  
Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, show_mean(iris) would print:  

function called show mean
for each column, calculates mean
```{r}
show_mean <- function(x) {
  show_mean_out <- 
}


```

