---
title: "week_6_problems"
author: "Veronica Thompson"
date: "5/25/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
library(tidyverse)
```

### Week 6 - Data Wrangling II  

R4DS Chapter 12 - Tidy Data   
Read 12.1 - 12.7  
12.2.1 Problems think about 1, complete 2 - 3  
12.3.3 Problems 1-4  
12.4.3 Problems 1-3  
12.6.1 Problems 3-4  
Joins tutorial (http://malooflab.phytonetworks.org/apps/joins-tutorial/)  

### Chapter 12  

> “Tidy datasets are all alike, but every messy dataset is messy in its own way.” –– Hadley Wickham  

#### 12.2.1 problems  
* 2 and 3 *  

Problem 2.  
Compute the rate for table2, and table4a + table4b.  

```{r}
#table 2
table2_cases = table2 %>%
  filter(type == "cases")
table2_pop = table2 %>%
  filter(type == "population")

table2_rate <- bind_cols(table2_cases, table2_pop) %>%
  mutate(case_rate = count/count1) %>%
  view()
```

```{r}
#table 4a and 4b

table4a$rate_1999 = table4a$`1999` / table4b$`1999`
table4a$rate_2000 = table4a$`2000` / table4b$`2000`
```

Problem 3.  
```{r}
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```

```{r}
table2 %>%
  filter(type == "cases") %>%
  ggplot(aes(year, count)) +
    geom_line(aes(group = country), color = "grey50") +
    geom_point(aes(color = country))
```

must filter cases data first!

#### 12.3.3 Problems 1-4  
Problem 1.  
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>%
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return", names_ptypes = list(year = factor())
               )
```

column names must be characters so they must be manually cooreced as numbers after pivot_longer. chr to nmber isn't allowed through names_ptypes.  

Problem 2.  
```{r}
table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
table4a
```

needed back ticks around the coumn names because numbers are a mess.  

Problem 3.  
```{r}
people <- tribble(
  ~name,             ~names,  ~values,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

There are two individuals named Phillip Woods here and they cannot be distinguished with the current data. Add and ID number to make this work..

```{r}
people <- tribble(
  ~name,             ~names,  ~values, ~ID_number,
  #-----------------|--------|--------|-----------
  "Phillip Woods",   "age",       45,  12234,
  "Phillip Woods",   "height",   186,  12234,
  "Phillip Woods",   "age",       50,  33324,
  "Jessica Cordero", "age",       37,  55555,
  "Jessica Cordero", "height",   156,  55555
)

#could also use mutate to assign ID numbers in a new column.  

people %>%
  pivot_wider(names_from = names, values_from = values) %>%
  print
```

Problem 4.  
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

#I want the columns to be sex, pregnant, not_pregnant

preg %>%
  pivot_longer(cols = c(male, female), names_to = "sex", values_to = "count") %>%
  pivot_wider(names_from = pregnant, values_from = count, names_prefix = "pregnant_")
```

#### 12.4.3 Problems 1-3  

Problem 1.  
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three")) #gives a warning

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge") #could also drop

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three")) #gives a warning

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left") #determines if NA leads or follows
```

Problem 2.  
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), remove = F, extra = "drop")
```

remove = F keeps the original column(s). set to T by default.  

Problem 3.  
extract is similar but uses regex to separate, can't think of an example right now.... maybe hypenated names? uniting is easier than separating, less information to define.  

#### 12.6.1 Problems 3-4  

Problem 3.  
```{r}
who %>%
  unite("codes", c(country, iso2, iso3)) %>%
  count(codes) %>%
  nrow()
  
who %>%
  count(country) %>%
  nrow()
```

Problem 4.  
For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.  
```{r}
who_tidy = who %>%
  pivot_longer(
    cols = new_sp_m014:newrel_f65, 
    names_to = "key", 
    values_to = "cases", 
    values_drop_na = TRUE
  ) %>% 
  mutate(
    key = stringr::str_replace(key, "newrel", "new_rel")
  ) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)

who_summary <- who_tidy %>%
  group_by(country, year, sex) %>%
  summarise(total_cases = sum(cases, na.rm = T)) 
```

This is a TON of data, let's filter for the ton 10 countries by case load.  
```{r}
who_summary_ranked <- who_summary %>%
  group_by(country) %>%
  summarise(country_cases = sum(total_cases)) %>%
  mutate(rank = percent_rank(country_cases)) %>%
  arrange(desc(country_cases)) 
```

let's take the top 10% of countries for the plot. Based on this I should filter for country cases > 310000

```{r}
who_summary_ranked = who_summary_ranked %>%
  filter(country_cases > 310000)
```


```{r}
who_summary %>%
  filter(country %in% who_summary_ranked$country) %>%
  ggplot(aes(year, total_cases, color = country)) +
  geom_point(position = position_jitter(), a = .4) +
  facet_wrap(sex~.) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

