---
title: "week2_rundergrad"
author: "Veronica Thompson"
date: "April 29, 2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Week of April 27  
R4DS Chapter 1 - Introduction to the book  
R4DS Chapter 2 - Introduction to Exploring data  
R4DS Chapter 3 - Data Visualization  
Read Sections 3.1 - 3.6 and 3.10 (3.7 - 3.9 are optional)  

### Complete the following problems  
3.2.4 Problems 1-5  
3.3.1 Problems 1-6  
3.5.1 Problems 1-6  
3.6.1 Problems 1-5 & select one graph from problem 6  

### 3.2.4 Problems 1 - 5  
#### 1
```{r}
ggplot(data = mpg)
```  

we get a blank plot, no aesthetics were defined.  
 
#### 2  
```{r}
nrow(mpg)
ncol(mpg)
```
234 rows and 11 columns  

#### 3
checked the help file. drv describes drive. There are 3 factors, front wheel, rear wheel, and 4wd.  
#### 4  
```{r}
ggplot(data = mpg) +
  geom_point(aes(x= cyl, y = hwy))
```

#### 5  
```{r}
ggplot(data = mpg) +
  geom_point(aes(x= class, y = drv))
```  
These are two  factors, so it doesn't show much but availible combinations.

### 3.3.1 Problems 1-6   
#### 1  
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```  
Color if defines in aes wll become a variable. We want to change the actual color to blue, not define the variable blue so it must be outside the aes mapping.  

#### 2  
```{r}
summary(mpg)
```

categorical seem to be listed as character vectors, we could parse this differently to figure it out.  

#### 3  
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy, size = year))
```  
can't map a continuous variable to shape  

#### 4
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy))
```  
It uses both aesthetics.  

#### 5  
```
vignette("ggplot2-specs")
```  
stroke is the outline size of each point. it definitely works with shape 21, probably only outlined shapes  

#### 6  
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ<5))
```  
labels by color given this argument.  


