---
title: "Week_4_questions"
author: "Veronica Thompson"
date: "May 12, 2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Week 4  
R4DS Chapter 7 - Exploratory Data Analysis  
Read Sections 7.1 - 7.8 
Complete the following problems 
* 7.3.4 Problems 1-4  
* 7.4.1 Problems 1-2  
* 7.5.1.1 Problems 1-3, 5, 6 (explore ggbeeswarm, but you do not have to describe methods)  

Read sections 7.5.2-7.5.3 and describe an experiment that would fall into each of these categories (two categorical variables and two continuous variables). No exercises required.

R4DS Chapter 8 - Workflow: Projects
Read Sections 8.1 - 8.5

#### 7.3.4 Problems 1-4  

Question 1.
```{r}
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = x), binwidth = 0.5)
```

again, some at zero for x variable...

```{r}
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = z), binwidth = 0.5)
```

```{r}
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = z), binwidth = 0.5) +
  coord_cartesian(ylim = 20)
```

```{r}
diamonds %>%
  filter(z >10 | z<2)
```

Question 2.  
```{r}
ggplot(diamonds) +
  geom_histogram(mapping = aes(price), binwidth = 10) +
  coord_cartesian( ylim = c(0, 100), xlim = c(0, 2000)
  )
```

Question 3.  
```{r}
diamonds %>%
  filter(carat == 1) %>%
  count()

diamonds %>%
  filter(carat == .99) %>%
  count()
```

you can't sell that rounded up crap for jewelery!

Question 4. 




#### 7.4.1 Problems 1-2   

Question 1.  
For bar plots, NA becomes it's own category. For histograms, NAs are removed becuase they are outside the definition of continuous variables.  

Question 2. 
```{r}
c(NA, 2:10, 2600) %>%
  mean()

c(NA, 2:10, 2600) %>%
  mean(na.rm  = TRUE)
````

####  7.5.1.1 Problems 1-3, 5, 6  

Question 1.  
```{r}
nycflights13::flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + sched_min / 60
  ) %>% 
  ggplot(mapping = aes(y = sched_dep_time, x = cancelled)) + 
    geom_boxplot(mapping = aes(fill = cancelled)) +
  coord_flip()
```

Question 2.  
```{r}
ggplot(diamonds, mapping = aes(cut, price)) +
  geom_jitter()
```

```{r}
ggplot(diamonds, mapping = aes(carat, price)) +
  geom_jitter()
```

```{r}
ggplot(diamonds, mapping = aes(cut, color)) +
         geom_count()
```

Let's stick with carat  
```{r}
ggplot(diamonds, mapping = aes(cut, carat)) +
  geom_boxplot()
```

carat trends negatively coorelated with cut  

Question 4.  
```{r}
library(lvplot)
ggplot(diamonds, mapping = aes(cut, carat)) +
  geom_lv()
```

WOW hate this. 