---
title: "FT 05_2020"
author: "Veronica Thompson"
date: "May 1, 2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library(RColorBrewer)
library("lme4")
library("lmerTest")
library(emmeans)
```

## Flowering Time  
### Field season 2019  
### analysis for Rundergrad club, spring 2020  

Outline:  
1. Set aesthecics  
2. Load data  
3. Plot  
4. Stat summary - coming soon!  

*Note*: when starting this document find/replace "FT" with appropriate experiment descriptor. Here I used "FT"  
Each coade chunk needs "r" indicator added as it is updated.  

### Input aesthectics  
Colors set to blue/green qualitative palette for now.  
```{r aesthetics}
#edit to rename genotypes on the plots
FT_name_key = tibble(
  genotype = c("WT", "elf3"),
  new_geno = c("GV 342", "elf3")
)

#set level and colors
FT_levels = c("GV 342", "elf3")
FT_colors = c("GV 342" = "#1f78b4", 
              "elf3" = "#33a02c")
```

 
### load data  
```{r load data}
#load  
FT_data <- read_csv("master_flowering_time_Data_2020_05_01.csv",
                    locale = locale(date_format = "%m_%d_%Y"),
                     col_types = cols(
                       DNA_ID = col_character(),
                       field_ID = col_character(),
                       genotype = col_factor(),
                       group = col_factor(),
                       row = col_factor(),
                       `end?` = col_factor(),
                       condition = col_factor(),
                       germ = col_date(),
                       yellow = col_date(),
                       flowering = col_date(),
                       final = col_date(),
                       diameter = col_double(),
                       direction = col_character(),
                       height = col_double(),
                       leaf_num = col_double(),
                       ax_heads = col_double(),
                       notes = col_character()))

#rename genotypes  
FT_data = left_join(FT_data, FT_name_key, by = "genotype")

#relevel genotypes  
FT_data$new_geno = factor(FT_data$new_geno, levels = FT_levels)

#removed the NAs in flowering. this is peice of data we are most interested in!
FT_data = FT_data %>%
  filter(flowering != is.na(flowering))

summary(FT_data)
```

### Calculate new variables  
	Flowering time (days between germ and flowering)  
	Flowering duration (days between flowering and final)  
	Flowering rate (diameter / duration)  
	Leaf frequency (height/ leaves)  
```{r}
FT_data_full = FT_data %>%
  mutate(
    yellow_time = yellow - germ,
    flowering_time = flowering - germ,
    final_time = final - germ,
    flowering_duration = final - flowering,
    flowering_rate = diameter / as.double(flowering_duration),
    leaf_freq = height / leaf_num
  )

#there are 4 NAs in new_geno, investigate

summary(FT_data_full)
```



### Plotting  
```{r plot_FT_facet}
ggplot(FT_data_full, 
       aes(new_geno, flowering_time, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = 0.5, position = position_jitter()) +
  facet_grid(.~ group) +
  labs(title = "Flowering time",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
        legend.position = c(.8,1.2)) + #LOVE this
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```
```{r}
FT_data_full %>%
  filter(new_geno == "GV 342") %>%
  filter(flowering_time > 60)
```

super interesting that there are so many value over 60. Not sure what is going on with planting A WT. 
```{r plot_FT_full}
ggplot(FT_data_full, 
       aes(new_geno,yellow_time, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(alpha = 0.5) +
  #facet_grid(.~ group) +
  labs(title = "Yellow petals time",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) +
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

```{r plot_FT_full}
ggplot(FT_data_full, 
       aes(new_geno, flowering_time, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(alpha = 0.5) +
  #facet_grid(.~ group) +
  labs(title = "Flowering time",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) +
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```
```{r plot_FT_full}
ggplot(FT_data_full, 
       aes(new_geno, final_time, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(alpha = 0.5) +
  #facet_grid(.~ group) +
  labs(title = "Final flowering time",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) +
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

```{r plot_duration_facet1}
ggplot(FT_data_full, 
       aes(new_geno, flowering_duration, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(alpha = 0.5) +
  facet_grid(.~ group) +
  labs(title = "Flowering time",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) +
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

What's with the baby group in planting A?

```{r}
FT_data_full %>%
  filter(flowering_duration < 2.5)
```

definitely an input error! oops! Let's filter first, then plot

```{r plot_duration_facet2}
FT_data_full %>%
  filter(flowering_duration > 1) %>%
ggplot(aes(new_geno, flowering_duration, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(alpha = 0.5) +
  facet_grid(.~ group) +
  labs(title = "Flowering duration",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) +
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

B WT is interesting. Wonder if it was cloudy one of those days?  

```{r plot_FT_full2}
FT_data_full %>%
  filter(flowering_duration > 1) %>%
ggplot(aes(new_geno, flowering_duration, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = 0.5, position = position_jitter()) +
  #facet_grid(.~ group) +
  labs(title = "Flowering duration",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.1)) + #my legend ran away on this one
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

snore.

For flowering rate we have the same filtering problems...
```{r plot_flower_rate_facet}
FT_data_full %>%
  filter(flowering_rate <5) %>%
ggplot(aes(new_geno, flowering_rate, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = 0.5, position = position_jitter()) +
  facet_grid(.~ group) +
  labs(title = "Flowering duration",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) + #
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

```{r plot_flower_rate_full}
FT_data_full %>%
  filter(flowering_rate <5) %>%
ggplot(aes(new_geno, flowering_rate, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_point(alpha = 0.5, position = position_jitter()) +
  #facet_grid(.~ group) +
  labs(title = "Flowering duration",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "days",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.1)) + #my legend ran away on this one
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

```{r plot_leaf_freq_facet1}
ggplot(FT_data_full, 
       aes(new_geno, leaf_freq, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(alpha = 0.5) +
  facet_grid(.~ group) +
  labs(title = "Flowering time",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "Average stem length/leaf (cm)",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) +
  #scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

```{r plot_leaf_freq_facet2}
ggplot(FT_data_full, 
       aes(new_geno, leaf_freq, 
           fill = new_geno, 
           shape = new_geno)) + 
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(alpha = 0.5) +
  #facet_grid(.~ group) +
  labs(title = "Flowering time",
       subtitle = "Summer 2019 field trial",
       x = "genotype",
       y = "Average stem length/leaf (cm)",
       fill = "Genotype",
       shape = "Genotype") +
  #theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) +
  #scale_y_continuous(breaks = c(2:5)) +
  #scale_color_manual(values = ) + 
  scale_fill_manual(values = FT_colors)
```

### Challenge - pivot table  
Make a plot using pivot table that they can recreate...  

```{r}
str(FT_data_full)
```


```{r FT_small}
FT_data_small = FT_data_full[c(18,4, 6, 7, 9:11)]
str(FT_data_small)  
```

```{r FT_gather}
FT_gather = FT_data_small %>%
  gather( 
       key = "dev_stage", 
       value = "dev_date",
       yellow, flowering, final)
head(FT_gather)
```

```{r}
ggplot(FT_gather, 
       aes(group, dev_date)) +
  geom_boxplot(aes(linetype = condition, 
                   fill = dev_stage),
               position = position_dodge2(padding = 0.2)) + #order matters!
  geom_point(aes(color = dev_stage, 
                 shape = new_geno),
                 alpha = 0.7, position = position_jitter()) +
  #facet_grid(.~ group) +
  labs(title = "Flowering trial Dates",
       subtitle = "Summer 2019 field trial",
       x = "date",
       y = "planting group") +
  theme(axis.text.x = element_text(angle = 65, vjust = .5)) +
  theme(legend.direction = "horizontal",
      legend.position = c(.8,1.2)) +
  #scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22)) +
  scale_color_brewer(palette = "Set1") +
  coord_flip()
```

### Troubleshooting  

Someone clearly has a date issue if they are -50 days to flowering...  
```
filter(FT_data_full, flowering_time < 30)
```

flowering is "2019_03_28", should be "2019_08_23" (probably??)

```
FT_data_full = mutate(FT_data_full, flowering = replace(flowering, flowering == "2019-03-28", "2019-08-23"))
```
