---
title: "week6_advanced"
author: "Veronica Thompson"
date: "5/25/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

### Week 6 exercises

##### Dull560 LUC data set  
Graph time and luminescence for each genotype  
* Make sure data is formatted properly  
* Calculate the average luminescence of each genotype at each time point  
* Scale luminescence of all genotypes  
  + Universal background subtraction so trough values are close to 0 (subtract the same thing from all genotypes)  
  + Within each genotype, scale luminescence from 1 to 0 (so the peak = 1)  
  + Try this using *grouped functions* (grouped mutates/summaries)  
  + Also try this using *mapping loops* (from purrr) for comparison (or for loops if you’d prefer)  
* Make your graph - think about aesthetics  
* Optional: calculate standard error of luminescence of each genotype at each time point and use that to create error bars  
  
```{r}
clock_data <- read_csv("Dull 560_All_Results.csv",
                     col_types = cols(
                       Genotype = col_factor()))

clock_data <- select(clock_data, 1:6, 19:89)

head(clock_data)
summary(clock_data$Genotype)
```

average lum, mins, and maxes at each timepoint by group
```{r}
clock_long <- clock_data %>%
  pivot_longer(cols = 7:77, names_to = "ZT", values_to = "lum")


clock_long$ZT <- sapply(clock_long$ZT, as.integer)

clock_ave_lum <- clock_long %>%
  group_by(Genotype, ZT) %>%
  summarise(average_lum = mean(lum, na.rm = T))

head(clock_ave_lum)
```

Next, subtract universal background -> what is the lowest minimum?  
```{r}
min(clock_long$lum)
```

subtract universal background from all lum values
```{r}
clock_scaled <- clock_long %>%
  mutate(lum_background = lum - min(clock_long$lum))

summary(clock_scaled$lum_background)
```

find max values to scale to by genotype
```{r}
clock_scaled_geno <- clock_scaled %>%
  group_by(Genotype) %>%
  summarise(scaled_min = min(lum_background),
            scaled_max = max(lum_background))

clock_scaled_geno
```

scale all values by genotype
```{r}
clock_scaled_final <- clock_scaled %>% 
  group_by(Genotype) %>%
  mutate(lum_scaled = lum_background / max(lum_background)) #this can't be right...

summary(clock_scaled_final$lum_scaled) 

```

```{r}
clock_scaled_final %>%
  sample_n(10) %>%
  filter(ZT > 48 & ZT <144) %>%
  ggplot(aes(ZT, lum_scaled, color = Genotype)) +
  geom_point(a=.2)
```

ah yes... sampling doesn't work for this format. make a random sample list to use for filtering. 

```{r}
clock_samples <- clock_data %>% 
  sample_n(50)

clock_scaled_final %>%
  filter(ZT >= 48 & ZT <= 144) %>%
  filter(Region %in% clock_samples$Region) %>%
  ggplot(aes(ZT, lum_scaled, color = Region, shape = Genotype)) +
  geom_point(alpha =.4) +
  geom_line(aes(group  = Region)) +
  facet_grid(Genotype ~.) +
  labs(title = "Scaled luminescence",
       subtitle = "Randomly sampled 8X contstruct",
       x = "ZT",
       y = "scaled lum",
       shape = "Genotype") + 
  scale_fill_manual() + 
  scale_x_continuous(breaks = c(48, 72, 96, 120, 144))
```

I'm supposed to be plotting averages....
scale the averaged by timepoint data
```{r}
clock_scaled_ave <- clock_ave_lum %>%
  mutate(ave_background = average_lum - min(average_lum)) %>%
  group_by(Genotype) %>%
  mutate(ave_scaled = ave_background / max(ave_background))

clock_scaled_ave %>%
  filter(ZT >= 48 & ZT <= 144) %>%
  ggplot(aes(ZT, ave_scaled, color = Genotype)) +
  geom_point(alpha =.4) +
  geom_line() +
  labs(title = "Scaled luminescence",
       subtitle = "Average values by genotype",
       x = "ZT",
       y = "scaled lum", 
       color = "Genotype") + 
  scale_fill_manual() + 
  scale_x_continuous(breaks = c(48, 72, 96, 120, 144))
```

BITE ME