---
title: "week5_problems"
author: "Veronica Thompson"
date: "May 15, 2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Week 5 - R4DS Data Wrangling I  
### Week of May 18  

R4DS Chapter 9 - Introduction  

R4DS Chapter 10 - Tibbles  
* Read 10.1 - 10.4  
* 10.5 Problems 1-2, 4-5  

R4DS Chapter 11 - Data Import 
* Read 11.1 - 11.6  
* 11.2.2 Problems 1, 4 - 5  
* 11.3.5 Problems 1-3, 7  

#### Chapter 10 questions  

Question 1.  
```{r}
print(mtcars)

as_tibble(mtcars) %>%
  print()
```

hmm, there's definitely not row names in a tibble.  

Question 2.  
```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x #called column xyz... weird
df[, "xyz"] 
df[, c("abc", "xyz")]
```

```{r}
tib <- tibble(abc = 1, xyz = "a")
tib$x #doesn't call anything
tib[, "xyz"] #easier to read here
tib[, c("abc", "xyz")] # no thanks...
```

Question 4.  
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

select(annoying, `1`)

ggplot(annoying, aes(x = `1`, `2`)) +
  geom_point()

ggplot(annoying, aes(x = annoying$`1`, y= annoying$`2`)) +
  geom_point()

annoying = annoying %>%
  mutate(`3` = `2` / `1`)

rename(annoying, 
       one = `1`,
       two = `2`, 
       three = `3`) %>%
  print()
```

Question 5.  
enframe makes a tibble.  
```{r}
#from help file examples
enframe(c(a = 5, b = 7))
enframe(list(one = 1, two = 2:3, three = 4:6)) #weird...
```

R4DS Chapter 11 - Data Import 
* Read 11.1 - 11.6  
* 11.2.2 Problems 1, 4 - 5  
* 11.3.5 Problems 1-3, 7  

#### Problems 11.2   
Question 1.  
```{r}
read_delim(delim = "|",
          "a|b|c\n1|2|.")
```

Question 4.  
```{r}
read_csv("x,y\n1,'a,b'",
         quote = "'")
```

Question 5.  
```{r}
read_csv("a,b\n1,2,3\n4,5,6") #missing a header name so last column is dropped
read_csv("a,b,c\n1,2\n1,2,3,4") #either new line out of place or misssing a col again. empty cells are filled with NA.
read_csv("a,b\n\"1") #I'm not sure what this is supposed to be...
read_csv("a,b\n1,2\na,b") #??? 
read_csv("a;b\n1;3") # i think semicolons should be colons?
```

#### Problems 11.3  
Question 1.  
all of the arguments seem important. For my work is is date_format.  

Question 2.  
```{}
locale(decimal_mark = ",", grouping_mark = ",")
Error: `decimal_mark` and `grouping_mark` must be different
```

```{r}
locale(decimal_mark = ",") #changes grouping to "."
```

Question 3.  
```{r}
parse_date("2019_03_17", locale = locale(date_format = "%Y_%m_%d"))
```

Question 7.  
```{r}
d1 <- parse_date("January 1, 2010", locale = locale(date_format = "%B %d, %Y"))
d2 <- parse_date("2015-Mar-07", locale = locale(date_format = "%Y-%b-%d"))
d3 <- parse_date("06-Jun-2017", locale = locale(date_format = "%d-%b-%Y"))
d4 <- parse_date(c("August 19 (2015)", "July 1 (2015)"), locale = locale(date_format = "%B %d (%Y)"))
d5 <- parse_date("12/30/14", locale = locale(date_format = "%m/%d/%y")) # Dec 30, 2014
(t1 <- parse_time("1705", locale = locale(time_format = "%H%M")))
(t2 <- parse_time("11:15:10.12 PM", locale = locale(time_format = "%I:%M:%OS %p")))
```

DONE.